---
title: "Joe-Jianing-STAT346HW3"
author: "Jianing Ren and Joe LaRocca"
date: "28/09/2020"
output: html_document
---

## 1
### a.
```{r}
set.seed(100)
smpl=matrix(rnorm(40000,mean=5,sd=3),nrow=10000,ncol=4)
unbias=apply(smpl,1,function(x) sum((x-5)^2))
mean(unbias)
```

The means of $\sigma^2$ is 35.9, which is roughly $n\sigma^2=36$

### b.
```{r}
bias=apply(smpl,1,function(x) sum((x-mean(x))^2))
mean(bias)
```

When we use the sample mean instead of using global mean hindsightly, we get 26.88, which is roughly equal to $(n-1)\sigma^2=27$. 

## 2

```{r}
set.seed(100)
x=rnorm(100,5,3)

y=sapply(x,function(x)4+3*x+rnorm(100))
ei=apply(y,1,function(y) sum((y-3*x-4)^2))
mean(ei)
```

Indeed, the average sum of squares for the 100 regressions is roughly 98, which is $(n-2)\sigma^2$. In order to get unbiased estimator for $\sigma^2$, divide $\sum{e_i^2}$ by n-2.

## 3 

To prove that $b_1$ is unbiased estimator of $B_1$, we need to show that $E(b_1)=B_1$.

$$
\begin{aligned}
E(b_1)&=\sum{k_iE(y_i)}\\
&=\frac{1}{s_{xx}}\sum(x_i-\bar{x})(\beta_0+\beta_1x_i)\\
&=\frac{1}{s_{xx}}\sum(x_i\beta_0+\beta_1x_i^2-\beta_0\bar{x}-\beta_1\bar{x}x_i)\\
&=\frac{1}{s_{xx}}[\beta_0\sum{(x_i-\bar{x})}+\beta_1\sum{x_i(x_i-\bar{x})}]\\
\end{aligned}
$$

Note that $\sum{(x_i-\bar{x})}=0$, 

and $\sum{x_i(x_i-\bar{x})}=\sum{x_i^2-\bar{x}\sum{x_i}}=\sum{x_i^2}-n\bar{x}^2=s_{xx}$. 

Therefore, we have $E(b_1)=\frac{1}{s_{xx}}[s_{xx}\beta_1]=\beta_1$. 

## 4
### ???a
```{r}
suppressMessages(library(mosaic))
bodyfat=read.delim("https://dasl.datadescription.com/download/data/3079")
set.seed(100)
res=do(1000)*coef(lm(Weight~Height,data=resample(bodyfat)))[2]
sd(res[,1])
```

### b
$$
\begin{aligned}
V(b_1)&=V(\sum{k_iy_i})\\
&=V(k_1y_1)+...+V(k_ny_n)~(the~y's~are~independent)\\
&=k_1^2V(y_1)+...+k_n^2V(y_n)\\
&=\sum{k_i^2}V(y_i)\\
&=\sigma^2\sum{k_i^2}\\
&=\sigma^2\frac{\sum{(x_i-\bar{x})^2}}{s_{xx}^2}\\
&=\sigma^2/s_{xx}\\
&=\frac{\sum{e_i^2}}{s_{xx}(n-2)}
\end{aligned}
$$

### ???c
```{r}
get_sxx=function(x)sum((x-mean(x))^2)
sxx=get_sxx(bodyfat$Height)
sse=sum(resid(lm(Weight~Height, data=bodyfat))^2)
sqrt(sse/sxx/(250-2))
```

## 4
### a

$$

$$
